---
title: "Predicting COVID-19 with Multiple Linear Regression"
author: "Scott Dolan"
date: "12/6/2020"
output: 
  html_document:
    code_folding: show
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)

# Clear plots
if(!is.null(dev.list())) dev.off()
# Clear console
cat("\014") 
# Clean workspace
rm(list=ls())

# Set Path
# Scott Path
path = '/home/scott/GIT/Modeling-COVID-19-with-R/'

# Zack Path
# path = ''

require("knitr")
opts_knit$set(root.dir = path)
```



### Load data and select desired desired data 
```{r}
rawdata <- read_csv('PA-COVID-19-Data-Merged.csv')


berks.dat <- rawdata %>% dplyr::filter(.,County == 'Berks' & Date >= '2020-06-01')
date_as_factor.berks <- data.frame(Day=c(1:length(unique(berks.dat$Date))),Date=unique(berks.dat$Date))
berks.dat <- berks.dat %>% left_join(.,date_as_factor.berks,by=c("Date"))
berks.model.dat <- berks.dat %>% dplyr::filter(.,Date <= '2020-12-15')
berks.verify.dat <- berks.dat %>% dplyr::filter(.,Date > '2020-12-15')


philly.dat <- rawdata %>% dplyr::filter(.,County == 'Philadelphia' & Date >= '2020-06-01')
date_as_factor.philly <- data.frame(Day=c(1:length(unique(philly.dat$Date))),Date=unique(philly.dat$Date))
philly.dat <- philly.dat %>% left_join(.,date_as_factor.philly,by=c("Date"))
philly.model.dat <- philly.dat %>% dplyr::filter(.,Date <= '2020-12-15')
philly.verify.dat <- philly.dat %>% dplyr::filter(.,Date > '2020-12-15')
```
### Berks County model 1 using linear regression and power transformation
```{r}
# DV  : # of New Cases
# IVs : Day

getMod = function(df){
  p = seq(0,4,0.01)
  Rsquare_p = c(0,0,0,0)
  p.value <- 1
  model = list(NULL)
  for(i in c(2:length(p))){

    modDat <- df %>%  dplyr::mutate(., Xdat = (Day^p[i]-1)/p[i]) %>% drop_na()
    fit <- modDat %>% lm(new.cases ~ Xdat,.)
    sumFit <- summary(fit)
    f <- sumFit$fstatistic
    p.value.fit <- pf(f[1],f[2],f[3],lower.tail=F)
    if(sumFit$adj.r.squared > Rsquare_p[2] & p.value.fit < p.value){
      Rsquare_p <- c(sumFit$r.squared,sumFit$adj.r.squared,p[i],i)
      model <- fit
      p.value <- p.value.fit
      model.sum <- sumFit
    }
  }
  if(is_empty(model)==TRUE){
      print('Search higher power')
      Rsquare_p <- c(sumFit$r.squared,sumFit$adj.r.squared,p[i],i)
      model <- fit
      p.value <- p.value.fit
      model.sum <- sumFit
  }

  return(list(model,Rsquare_p,modDat$Day,p.value,model.sum))
}
berks.model1 <- getMod(berks.model.dat)
```

#### Explore model 1 for Berks County using linear regression (unweighted)
```{r}
summary(berks.model1[[1]])
model1 = berks.model1[[1]]
modelDat <- data.frame(model1$model)
power <- berks.model1[[2]][[3]]
plot(modelDat$Xdat,modelDat$new.cases,main = 'Model 1 for Berks County PA')
abline(model1)

# get summary and plots of fit
layout(matrix(1:4,2,2))
plot(model1) 
par(mfrow=c(1,1))

# Make predictions
Yhat <- function(B0,B1,day,p){B0+B1*(day^p-1)/p}

y.int <- model1$coefficients[1]
slop <- model1$coefficients[2]
berks.prediction <- Yhat(y.int,slop,berks.verify.dat$Day,power)
plot((berks.dat$Day^power-1)/power , berks.dat$new.cases,main = 'Berks Model1 with Predictions')
prediction.days <- (berks.verify.dat$Day^power-1)/power
prediction <- Yhat(y.int,slop,berks.verify.dat$Day,power)
points(prediction.days, prediction, pch=16,col='red')
sprintf('Acutal new case count for %s was %.2f',berks.verify.dat$Date[1],berks.verify.dat$new.cases[1])
sprintf('Predicted new case count for %s was %.2f',berks.verify.dat$Date[1],prediction[1])
sprintf('Acutal new case count for %s was %.2f',berks.verify.dat$Date[2],berks.verify.dat$new.cases[2])
sprintf('Predicted new case count for %s was %.2f',berks.verify.dat$Date[2],prediction[2])
sprintf('Acutal new case count for %s was %.2f',berks.verify.dat$Date[3],berks.verify.dat$new.cases[3])
sprintf('Predicted new case count for %s was %.2f',berks.verify.dat$Date[3],prediction[3])
```

### Berks County model 2 with non-linear regression (weighted)
```{r}
residMod1 <- abs(resid(model1))
model2=lm(residMod1~modelDat$Xdat)
fitted=fitted(model2)

weight=(1/(fitted*fitted))

wls=lm(modelDat$new.cases~modelDat$Xdat, weights=weight)
summary(wls)

model2.berks <- wls
summary(model2.berks)
layout(matrix(1:4,2,2))
plot(model2.berks) 
par(mfrow=c(1,1))
model2 <- model2.berks$model
plot(model2$`modelDat$Xdat`,model2$`modelDat$new.cases`)
abline(model2.berks)
y.int2 <- model2.berks$coefficients[1]
slop2 <- model2.berks$coefficients[2]
berks.prediction <- Yhat(y.int,slop,berks.verify.dat$Day,power)
plot((berks.dat$Day^power-1)/power , berks.dat$new.cases,main = 'Berks Model1 with Predictions')
prediction.days <- (berks.verify.dat$Day^power-1)/power
prediction <- Yhat(y.int2,slop2,berks.verify.dat$Day,power)
points(prediction.days, prediction, pch=16,col='red')
sprintf('Acutal new case count for %s was %.2f',berks.verify.dat$Date[1],berks.verify.dat$new.cases[1])
sprintf('Predicted new case count for %s was %.2f',berks.verify.dat$Date[1],prediction[1])
sprintf('Acutal new case count for %s was %.2f',berks.verify.dat$Date[2],berks.verify.dat$new.cases[2])
sprintf('Predicted new case count for %s was %.2f',berks.verify.dat$Date[2],prediction[2])
sprintf('Acutal new case count for %s was %.2f',berks.verify.dat$Date[3],berks.verify.dat$new.cases[3])
sprintf('Predicted new case count for %s was %.2f',berks.verify.dat$Date[3],prediction[3])
```

### Philadelphia County model 1 using linear regression and power transformation
```{r}
# DV  : # of New Cases
# IVs : Day

getMod = function(df){
  p = seq(0,10,0.01)
  Rsquare_p = c(0,0,0,0)
  p.value <- 1
  model = list(NULL)
  for(i in c(2:length(p))){

    modDat <- df %>%  dplyr::mutate(., Xdat = (Day^p[i]-1)/p[i]) %>% drop_na()
    fit <- modDat %>% lm(new.cases ~ Xdat,.)
    sumFit <- summary(fit)
    f <- sumFit$fstatistic
    p.value.fit <- pf(f[1],f[2],f[3],lower.tail=F)
    if(sumFit$adj.r.squared > Rsquare_p[2] & p.value.fit < p.value){
      Rsquare_p <- c(sumFit$r.squared,sumFit$adj.r.squared,p[i],i)
      model <- fit
      p.value <- p.value.fit
      model.sum <- sumFit
    }
  }
  if(is_empty(model)==TRUE){
      print('Search higher power')
      Rsquare_p <- c(sumFit$r.squared,sumFit$adj.r.squared,p[i],i)
      model <- fit
      p.value <- p.value.fit
      model.sum <- sumFit
  }

  return(list(model,Rsquare_p,modDat$Day,p.value,model.sum))
}
philly.model1 <- getMod(philly.model.dat)
```

#### Explore model 1 for Philadelphia County using linear regression (unweighted)
```{r}
summary(philly.model1[[1]])
model1 = philly.model1[[1]]
modelDat <- data.frame(model1$model)
power <- philly.model1[[2]][[3]]
plot(modelDat$Xdat,modelDat$new.cases,main = 'Model 1 for Philadelphia County PA')
abline(model1)

# get summary and plots of fit
layout(matrix(1:4,2,2))
plot(model1) 
par(mfrow=c(1,1))

# Make predictions
Yhat <- function(B0,B1,day,p){B0+B1*(day^p-1)/p}

y.int <- model1$coefficients[1]
slop <- model1$coefficients[2]
philly.prediction <- Yhat(y.int,slop,philly.verify.dat$Day,power)
plot((philly.dat$Day^power-1)/power , philly.dat$new.cases,main = 'Philadelphia Model1 with Predictions')
prediction.days <- (philly.verify.dat$Day^power-1)/power
prediction <- Yhat(y.int,slop,philly.verify.dat$Day,power)
points(prediction.days, prediction, pch=16,col='red')
sprintf('Acutal new case count for %s was %.2f',philly.verify.dat$Date[1],philly.verify.dat$new.cases[1])
sprintf('Predicted new case count for %s was %.2f',philly.verify.dat$Date[1],prediction[1])
sprintf('Acutal new case count for %s was %.2f',philly.verify.dat$Date[2],philly.verify.dat$new.cases[2])
sprintf('Predicted new case count for %s was %.2f',philly.verify.dat$Date[2],prediction[2])
sprintf('Acutal new case count for %s was %.2f',philly.verify.dat$Date[3],philly.verify.dat$new.cases[3])
sprintf('Predicted new case count for %s was %.2f',philly.verify.dat$Date[3],prediction[3])
```


### Create model for entire state of Pennsylvania
```{r}
# To model the state, parameters need to be combined for a state total for each day
start.date <- '2020-06-01'
end.date <- '2020-12-15'

PA.dat <- rawdata %>%   dplyr::filter(Date >= as.Date(start.date)) %>% drop_na() 
Pa.date_as_factor <- data.frame(Day=c(1:length(unique(PA.dat$Date))),Date=unique(PA.dat$Date))
PA.dat <- PA.dat %>% left_join(.,Pa.date_as_factor,by=c("Date")) %>% group_by(.,Day) %>% dplyr::mutate(.,Date=unique(Date),Day= Day,new.cases.PA = sum(new.cases), daily.death.count.PA = sum(daily.death.count), patients.hospitalized.PA = sum(patients.hospitalized), patients.on.ventilators.PA = sum(patients.on.ventilators), population.2018.PA = sum(population.2018), airborn.isolation.beds.PA = sum(airborn.isolation.beds), adult.ICU.beds.PA = sum(adult.ICU.beds)) %>% dplyr::select(.,Date,new.cases.PA, daily.death.count.PA, patients.hospitalized.PA, Day, patients.on.ventilators.PA, airborn.isolation.beds.PA, adult.ICU.beds.PA) %>% unique(.) %>% ungroup(.) 

PA.model.dat <- PA.dat %>% dplyr::filter(Date <= as.Date(end.date))
PA.verify.dat <- PA.dat %>% dplyr::filter(Date > as.Date(end.date))

```

```{r}
# Time series model by Day # 
# DV = # of new cases per day
# IVS : Day 

getMod = function(df){
  p = seq(0,10,.1)
  Rsquare_p = c(0,0,0,0)
  p.value <- 1
  model = list(NULL)
  for(i in c(2:length(p))){

    modDat <- df %>%  dplyr::mutate(., Xdat = (Day^p[i]-1)/p[i]) %>% drop_na()
    #print(head(modDat))
    fit <- lm(new.cases.PA ~ Xdat, data = modDat)
    sumFit <- summary(fit)
    f <- sumFit$fstatistic
    p.value.fit <- pf(f[1],f[2],f[3],lower.tail=F)
    if(sumFit$adj.r.squared > Rsquare_p[2] & p.value.fit < p.value){
      Rsquare_p <- c(sumFit$r.squared,sumFit$adj.r.squared,p[i],i)
      model <- fit
      p.value <- p.value.fit
      model.sum <- sumFit
    }
  }
  if(is_empty(model)==TRUE){
      print('Search higher power')
      Rsquare_p <- c(sumFit$r.squared,sumFit$adj.r.squared,p[i],i)
      model <- fit
      p.value <- p.value.fit
      model.sum <- sumFit
  }

  return(list(model,Rsquare_p,modDat$Day,p.value,model.sum))
}

PA.model1 <- getMod(PA.model.dat)

# PA.model <- lm(new.cases.PA ~ Day + daily.death.count.PA + patients.hospitalized.PA + patients.on.ventilators.PA + airborn.isolation.beds.PA + adult.ICU.beds.PA + percent.Democrat.PA + percent.Republican.PA, data = PA.model.dat)

```

#### Explore model 1 for PA
```{r}
summary(PA.model1[[1]])
model1 = PA.model1[[1]]
modelDat <- data.frame(model1$model)
power <- PA.model1[[2]][[3]]
plot(modelDat$Xdat,modelDat$new.cases,main = 'Model 1 for Berks County PA')
abline(model1)
# get summary and plots of fit
layout(matrix(1:4,2,2))
plot(model1) 
par(mfrow=c(1,1))

# Make predictions
Yhat <- function(B0,B1,day,p){B0+B1*(day^p-1)/p}

y.int <- model1$coefficients[1]
slop <- model1$coefficients[2]

plot((PA.dat$Day^power-1)/power , PA.dat$new.cases.PA,main = 'PA Model 1 with Predictions')
prediction.days <- (PA.verify.dat$Day^power-1)/power
PA.prediction <- Yhat(y.int,slop,PA.verify.dat$Day,power)
points(prediction.days, PA.prediction, pch=16,col='red')
sprintf('Acutal new case count for %s was %.2f',PA.verify.dat$Date[1],PA.verify.dat$new.cases.PA[1])
sprintf('Predicted new case count for %s was %.2f',PA.verify.dat$Date[1],PA.prediction[1])
sprintf('Acutal new case count for %s was %.2f',PA.verify.dat$Date[2],PA.verify.dat$new.cases.PA[2])
sprintf('Predicted new case count for %s was %.2f',PA.verify.dat$Date[2],PA.prediction[2])
sprintf('Acutal new case count for %s was %.2f',PA.verify.dat$Date[3],PA.verify.dat$new.cases.PA[3])
sprintf('Predicted new case count for %s was %.2f',PA.verify.dat$Date[3],PA.prediction[3])
```


### Create weighted model for entire state of Pennsylvania
```{r}
summary(PA.model1[[1]])
model1 = PA.model1[[1]]
residMod1 <- abs(resid(model1))
model2=lm(residMod1~modelDat$Xdat)
fitted=fitted(model2)

weight=(1/(fitted*fitted))

wls=lm(modelDat$new.cases~modelDat$Xdat, weights=weight)
summary(wls)
#plot(wls)
plot(modelDat$Xdat,modelDat$new.cases.PA)
abline(wls)

power <- PA.model1[[2]][[3]]
# get summary and plots of fit
layout(matrix(1:4,2,2))
plot(wls) 
par(mfrow=c(1,1))
# 
# # Make predictions
# Yhat <- function(B0,B1,day,p){B0+B1*(day^p-1)/p}
# 
y.int <- wls$coefficients[1]
slop <- wls$coefficients[2]
# 
plot((PA.dat$Day^power-1)/power , PA.dat$new.cases.PA,main = 'PA WLS Model with Predictions')
prediction.days <- (PA.verify.dat$Day^power-1)/power
PA.prediction.wls <- Yhat(y.int,slop,PA.verify.dat$Day,power)
points(prediction.days, PA.prediction.wls, pch=16,col='red')
sprintf('Acutal new case count for %s was %.2f',PA.verify.dat$Date[1],PA.verify.dat$new.cases.PA[1])
sprintf('Predicted new case count for %s was %.2f',PA.verify.dat$Date[1],PA.prediction.wls[1])
sprintf('Acutal new case count for %s was %.2f',PA.verify.dat$Date[2],PA.verify.dat$new.cases.PA[2])
sprintf('Predicted new case count for %s was %.2f',PA.verify.dat$Date[2],PA.prediction.wls[2])
sprintf('Acutal new case count for %s was %.2f',PA.verify.dat$Date[3],PA.verify.dat$new.cases.PA[3])
sprintf('Predicted new case count for %s was %.2f',PA.verify.dat$Date[3],PA.prediction.wls[3])
```